\chapter{Introduction}\label{introduction}
\input{master_thesis_content/introduction/introduction}
\chapter{Background}
\section{Related work}
\input{master_thesis_content/background/background_relatedwork}
\input{master_thesis_content/background/background_climatenet}
\input{master_thesis_content/background/background_sam}
\input{master_thesis_content/background/background_metrics}

\chapter{Approach}
\input{master_thesis_content/approach/approach_pipeline}
\section{Architectures}
\label{sec:architectures}

This section describes the various architectural components utilized in the development of ClimateSAM. To adapt the SAM for ClimateNet Dataset, the framework incorporates specialized modules for data adaptation, feature refinement, and autonomous guidance. The overall architecture, as visualized in Figure \ref{fig:pipeline}, is designed to handle the 16-channel atmospheric grids of ClimateNet while transitioning SAM toward a prompt-free operational state.

The system comprises three main architectural pillars:
\begin{itemize}
    \item \textbf{Input Adapter}: A newly suggested ClimateInputAdapter is implemented as a learnable bottleneck convolutional module. It compresses the high-dimensional atmospheric data into a 3-channel pseudo-RGB representation, ensuring compatibility with the SAM image encoder while preserving critical physical signals.
    \item \textbf{Encoder Adapters}: Two Parameter-Efficient Fine-Tuning (PEFT) strategies are explored to adapt the frozen ViT backbone. These include a \textit{learnable token-based} mechanism to steer the model using latent representations and a \textit{LoRA-based} approach that utilizes low-rank matrix decomposition to update the query and value projections.
    \item \textbf{Prompt Generators}: To achieve autonomous detection, the pipeline evaluates two prompter architectures. We utilize \textit{CGNet} an established lightweight convolutional network for semantic segmentation alongside a newly introduced \textit{Multi-scale Fusion Module}. This module leverages intermediate encoder features to generate auxiliary masks, which are then converted into sparse geometric prompts for the mask decoder.
\end{itemize}

\input{master_thesis_content/approach/approach_model}
% \section{Hyperparameter Optimisation}
\input{master_thesis_content/approach/approach_dataset_prep}
\input{master_thesis_content/approach/approach_loss_function.tex}
\input{master_thesis_content/approach/approach_label_smooth.tex}
\section{Metrics}

\chapter{Result}
\section{Preliminary Result}
\input{master_thesis_content/results/result_preliminary}
\section{Adapt SAM to ClimateNet}
\input{master_thesis_content/results/result_lora}
\input{master_thesis_content/results/result_token}
\section{Automatic Prompt Generator for SAM}
\input{master_thesis_content/results/result_cgnet}
\input{master_thesis_content/results/result_multiscaled}

\chapter{Discussion}

% % \section{Motivation}
% % \section{Research Questions}
% \chapter{Approach}\label{approach}
% \input{Approach}



% \section{Implementation plan and schedules}
% \chapter{}

% 1. Topic introduction:

% The automated detection and tracking of two-dimensional (2-D) and three-dimensional (3-D) atmospheric features, such as cyclones, fronts, jet streams, and atmospheric rivers (ARs), in both simulation and observational data, has become integral to meteorology. These automated techniques find applications in weather forecasting, statistical and climatological studies, and visual data analysis.

% Traditionally, atmospheric features have been detected using physical and mathematical rules. Cyclones, for example, are identified through minima or maxima in variables like sea level pressure and lower-tropospheric vorticity, while fronts are detected via derivatives of thermal variables and threshold-based filters. Atmospheric rivers are identified based on thresholds in integrated water vapor transport. 



% motivation: why detecting tropical cyclone and atmospherical rivers important. (3 Sentences - with references)  traditionally how it was doned. Then there was  an interface and describe how climate scientist do it by hand and costs many hours.



% Example:
% The automated detection and tracking of 2-D and 3-D atmospheric features including cyclones, fronts, jet streams, or atmospheric rivers (ARs) in simulation and observation data has multiple applications in meteorology. For example, automatically detected features are used for weather forecasting (e.g., Hewson and Titley, 2010; Mittermaier et al., 2016; Hengstebeck et al., 2018), statistical and climatological studies (e.g., Dawe and Austin, 2012; Pena-Ortiz et al., 2013;
% Schemm et al., 2015; Sprenger et al., 2017; Lawrence and Manney, 2018), and visual data analysis (e.g., Rautenhaus et al., 2018; Bösiger et al., 2022; Beckert et al., 2023). Features are typically detected based on a set of physical and mathematical rules. For example, cyclones can be identified by searching for minima or maxima in variables including mean 
% sea level pressure and lower-tropospheric vorticity (Neu et
% al., 2013; Bourdin et al., 2022). Atmospheric fronts can be
% identified by means of derivatives of a thermal variable combined with threshold-based filters (Jenkner et al., 2010; Hewson and Titley, 2010; Beckert et al., 2023) and ARs based on thresholding and geometric requirements (Guan and Waliser,
% 2015; Shields et al., 2018).


% Then there is machine learning approach, startrted 

% the manual works in reading 





% 1. Short Motivation. (half page)
% 2. Related Works.
% - SAM
% - CAT-SAM
% - SOS
% - Tim Radke Work: 

% approach:


% Mask prompts:


% - Using a U-NET for dimension reductions.
% - 
