\chapter{Results and Analysis}\label{results}
This chapter is dedicated to presenting the results of our experiments. We will begin by showcasing the outcomes of the hyperparameter tuning process for the evaluation models and the augmentation techniques. Then we will present the results of both the classification and detection tasks for the basic augmentations. Finally, we will dive into the results of using GAN to enhance the performance of edge cases. To improve the visual representation, a smoothing average window of 0.7 was applied to all graphics.

\section{Baseline and Hyperparameter tuning}
In this section, we present the baseline for the YOLOv5s and EfficientNet for the given task, along with the optimal hyperparameters for the augmentation techniques. The baseline experiments were conducted without any data augmentation and involved training the models both with and without pre-trained weights. The evaluation metrics used to measure the performance of the models include F1 score Top-3 and mAP@[0.5:0.95].
\input{5baseline.tex}

\section{Results of basic augmentation techniques}
In this section, we present the outcomes of applying fundamental augmentations, which include Geometric, Color-Space, Cutout, and RandAugment techniques. The section is divided into two parts, with the first one showcasing the results of the detection task using YOLOv5s, and the second one presenting the findings of the classification task using EfficientNet B0.
\subsection{Detection Task}
\input{5resultdetection.tex}
\subsection{Classification Task}
\input{5resultclassification.tex}
\section{GAN-generated augmentations}
\input{5resultgan.tex}
